Faithful Variable Screening for\\ High-Dimensional Convex Regression

### 2014-09-16 tex/scam/experiments.tex
added section label
new plots
minor edits

\section{Experiments}
%\subsection{Simulations}
We first illustrate our methods using a simulation of the following regression problem
\begin{equation}\nonumber
         y_i = \bds{x}_{iS}^{\top}\bds{Q}\bds{x}_{iS} + \epsilon_i \quad (i=1,2,\ldots,n).
\end{equation}
Here $\bds{x}_{i}$ denotes data sample $i$ drawn from $\mathcal{N}(\bds{0},\bds{I}_{p})$, 
$\bds{x}_{iS}$ is a subset of $\bds{x}_i$ with dimension $|S|=5$, where $S$ represents the active feature set, and 
$\epsilon_i$ is the additive noise drawn from $\mathcal{N}(0,1)$. 
$\bds{Q}$ is a symmetric positive definite matrix of dimension $|S|\times{}|S|$. 
Notice that if $\bds{Q}$ is diagonal, then the true function is convex additive; otherwise the true function is convex but not additive.
For all the simulations in this section, we set $\lambda=4\sqrt{{\log(np)}/{n}}$.

In the first simulation, we set $\bds{Q}=\bds{I}_{|S|}$ (the additive
case), and choose $n=100, 200,\ldots,1000$ and $p=64,128,256,512$.
For each $(n,p)$ combination, we generate $200$ independent data
sets. For each data set we use SCAM to infer the model parameterized
by $\bds{h}$ and $\bds{\beta}$; see equation \eqref{np}. If
$\|\bds{\beta}_{k\cdot}\|_{\infty}<10^{-8}\ (\forall k\not\in{}S)$ and
$\|\bds{\beta}_{k\cdot}\|_{\infty}>10^{-8}\ (\forall k\in{}S)$, then
we declare correct support recovery. We then plot the probability of
support recovery over the $200$ data sets in Figure \ref{Support}(a).  We
observe that SCAM performs consistent variable selection when the true
function is convex additive.  
To give the reader a
sense of the running speed, the code runs in about $2$ minutes on one
data set with $n=1000$ and $p=512$, on a MacBook with 2.3 GHz Intel
Core i5 CPU and 4 GB memory.

In the second simulation, we study the case in which the true function
is convex but not additive. We generate four $\bds{Q}$ matrices
plotted in Figure \ref{Support}(b), where the diagonal elements are all $1$ and
the off-diagonal elements are $0.5$ with probability $\alpha$
($\alpha=0,0.2,0.5,1$ for the four cases). We fix $p=128$ and choose
$n=100,200,\ldots,1000$. We again run the SCAM optimization on $200$
independently generated data sets and plot the probability of recovery
in Figure \ref{Support}(c). The results demonstrate that SCAM performs
consistent variable selection even if the true function is not additive (but
still convex).

In the third simulation, we study the case of correlated design, where
$\bds{x}_{i}$ is drawn from $\mathcal{N}(\bds{0},\bds{\Sigma})$
instead of $\mathcal{N}(\bds{0},\bds{I}_{p})$, with
$\Sigma_{ij}=\nu^{|i-j|}$. We use the non-additive $\bds{Q}$ with
$\alpha=0.5$ and fix $p=128$.  The recovery curves for $\nu=02, 0.4,
0.6, 0.8$ are depicted in Figure \ref{Support}(d). As can be seen, for
design of moderate correlation, SCAM can still select relevant
variables well.

%\begin{figure}[!htpb]
%        \centering
%        \begin{subfigure}[b]{0.45\textwidth}
%                \centering
%                \includegraphics[width=\textwidth]{figs/Curve1}
%                 \caption{Additive model.}
%                \label{Curve1}
%        \end{subfigure}
%        \begin{subfigure}[b]{0.45\textwidth}
%                \centering
%                \includegraphics[width=\textwidth]{figs/Q}
%                \caption{Four $\bds{Q}$ matrices.}
%                \label{Q}
%        \end{subfigure}\\
%        \begin{subfigure}[b]{0.45\textwidth}
%                \centering
%                \includegraphics[width=\textwidth]{figs/Curve2}
%                \caption{Additive and non-additive models.}
%                \label{Curve2}
%        \end{subfigure}
%        \begin{subfigure}[b]{0.45\textwidth}
%                \centering
%                \includegraphics[width=\textwidth]{figs/Curve3}
%                \caption{Correlated design for non-additive model.}
%                \label{Curve3}
%        \end{subfigure}
%        \caption{Result of support recovery.}\label{Support}
%\end{figure}

%\begin{figure}[ht]
%\begin{center}
%\begin{tabular}{cccc}
%\hskip-20pt
%\includegraphics[width=.27\textwidth]{figs/Curve1} &
%\hskip-17pt
%\includegraphics[width=.27\textwidth]{figs/Q} &
%\hskip-17pt
%\includegraphics[width=.27\textwidth]{figs/Curve2} &
%\hskip-17pt
%\includegraphics[width=.27\textwidth]{figs/Curve3}  \\
%(a) additive model & (b) four $\bds{Q}$ matrices &
%(c) non-additive models & (d) correlated design
%\end{tabular}
%\end{center}
%\caption{Result of support recovery.}\label{Support}
%\end{figure}

@@

\begin{figure*}[!t]
\begin{center}
\begin{tabular}{cccc}
\hskip-20pt
\includegraphics[width=.29\textwidth]{figs/Curve1} &
\hskip-20pt
\includegraphics[width=.29\textwidth]{figs/Q} &
\hskip-20pt
\includegraphics[width=.29\textwidth]{figs/Curve2} &
\hskip-20pt
\includegraphics[width=.29\textwidth]{figs/Curve3}  \\
\hskip-10pt (a) additive model & 
\hskip-10pt (b) four $\bds{Q}$ matrices &
\hskip-10pt (c) non-additive models & 
\hskip-10pt (d) correlated design
\end{tabular}
\end{center}
\caption{Support recovery results where the additive assumption is
  correct (a), incorrect (b), (c), and with correlated design (d).}\label{Support}
\vskip10pt

>>>>

\def\x{x}
\def\Q{Q}
\def\bds#1{#1}

\section{Experiments}
\label{sec:thesims}
%\subsection{Simulations}
We first illustrate our methods using a simulation of the model
\begin{equation}\nonumber
         Y_i = \x_{iS}^{\top}\Q\x_{iS} + \epsilon_i \quad (i=1,2,\ldots,n).
\end{equation}
Here $\x_{i}$ denotes data sample $i$ drawn from $\mathcal{N}(\bds{0},\bds{I}_{p})$, 
$\x_{iS}$ is a subset of $\x_i$ with dimension $|S|=5$, where $S$
represents the relevant variable set, and 
$\epsilon_i$ is additive noise drawn from $\mathcal{N}(0,1)$. 
The matrix $\Q$ is a symmetric positive definite matrix of dimension $|S|\times{}|S|$. 
Note that if $\Q$ is diagonal, then the true function is convex
additive; 
otherwise the true function is convex but not additive.
For all simulations in this section, we set $\lambda=4\sqrt{{\log(np)}/{n}}$.

In the first simulation, we set $\Q=\bds{I}_{|S|}$ (the additive
case), and choose $n=100, 200,\ldots,1000$ and $p=64,128,256,512$.
For each $(n,p)$ combination, we generate $200$ independent data
sets. For each data set we use SCAM to infer the model parameterized
by $\bds{h}$ and $\bds{\beta}$; see equation \eqref{np}. If
$\|\bds{\beta}_{k\cdot}\|_{\infty}<10^{-8}\ (\forall k\not\in{}S)$ and
$\|\bds{\beta}_{k\cdot}\|_{\infty}>10^{-8}\ (\forall k\in{}S)$, then
we declare correct support recovery. We then plot the probability of
support recovery over the $200$ data sets in Figure \ref{Support}(a).  We
observe that SCAM performs consistent variable selection when the true
function is convex additive.  
To give the reader a
sense of the running speed, the code runs in about $2$ minutes on one
data set with $n=1000$ and $p=512$, on a MacBook with 2.3 GHz Intel
Core i5 CPU and 4 GB memory.

In the second simulation, we study the case in which the true function
is convex but not additive. We generate four $\Q$ matrices
plotted in Figure \ref{Support}(b), where the diagonal elements are all $1$ and
the off-diagonal elements are $0.5$ with probability $\alpha$
($\alpha=0,0.2,0.5,1$ for the four cases). We fix $p=128$ and choose
$n=100,200,\ldots,1000$. We again run the SCAM optimization on $200$
independently generated data sets and plot the probability of recovery
in Figure \ref{Support}(c). The results demonstrate that SCAM performs
consistent variable selection even if the true function is not additive (but
still convex).

In the third simulation, we study the case of correlated design, where
$\x_{i}$ is drawn from $\mathcal{N}(\bds{0},\bds{\Sigma})$
instead of $\mathcal{N}(\bds{0},\bds{I}_{p})$, with
$\Sigma_{ij}=\nu^{|i-j|}$. We use the non-additive $\Q$ with
$\alpha=0.5$ and fix $p=128$.  The recovery curves for $\nu=02, 0.4,
0.6, 0.8$ are depicted in Figure \ref{Support}(d). As can be seen, for
design of moderate correlation, SCAM can still select relevant
variables well.

%\begin{figure}[!htpb]
%        \centering
%        \begin{subfigure}[b]{0.45\textwidth}
%                \centering
%                \includegraphics[width=\textwidth]{figs/Curve1}
%                 \caption{Additive model.}
%                \label{Curve1}
%        \end{subfigure}
%        \begin{subfigure}[b]{0.45\textwidth}
%                \centering
%                \includegraphics[width=\textwidth]{figs/Q}
%                \caption{Four $\Q$ matrices.}
%                \label{Q}
%        \end{subfigure}\\
%        \begin{subfigure}[b]{0.45\textwidth}
%                \centering
%                \includegraphics[width=\textwidth]{figs/Curve2}
%                \caption{Additive and non-additive models.}
%                \label{Curve2}
%        \end{subfigure}
%        \begin{subfigure}[b]{0.45\textwidth}
%                \centering
%                \includegraphics[width=\textwidth]{figs/Curve3}
%                \caption{Correlated design for non-additive model.}
%                \label{Curve3}
%        \end{subfigure}
%        \caption{Result of support recovery.}\label{Support}
%\end{figure}

%\begin{figure}[ht]
%\begin{center}
%\begin{tabular}{cccc}
%\hskip-20pt
%\includegraphics[width=.27\textwidth]{figs/Curve1} &
%\hskip-17pt
%\includegraphics[width=.27\textwidth]{figs/Q} &
%\hskip-17pt
%\includegraphics[width=.27\textwidth]{figs/Curve2} &
%\hskip-17pt
%\includegraphics[width=.27\textwidth]{figs/Curve3}  \\
%(a) additive model & (b) four $\Q$ matrices &
%(c) non-additive models & (d) correlated design
%\end{tabular}
%\end{center}
%\caption{Result of support recovery.}\label{Support}
%\end{figure}

@@

\begin{figure*}[!t]
\begin{center}
\begin{tabular}{cccc}
\hskip-10pt
\includegraphics[width=.26\textwidth]{figs/Curve1} &
\hskip-10pt
\includegraphics[width=.26\textwidth]{figs/Q} &
\hskip-10pt
\includegraphics[width=.26\textwidth]{figs/Curve2} &
\hskip-10pt
\includegraphics[width=.26\textwidth]{figs/Curve4}  \\
\hskip-10pt (a) additive model & 
\hskip-10pt (b) four $\Q$ matrices &
\hskip-10pt (c) non-additive models & 
\hskip-10pt (d) correlated design
\end{tabular}
\end{center}
\caption{Support recovery results where the additive assumption is
  correct (a), incorrect (b), (c), and with correlated design (d).}\label{Support}
\vskip10pt
